\ripple: Differentiable \dfm{and Hardware-Accelerated} Waveforms for Gravitational Wave Data Analysis

### 2023-01-20 src/tex/ms.tex
Testing cache issue from yesterday
Adding edit

The discovery of gravitational waves (GWs)~\cite{LIGOScientific:2016aoc} from inspiraling and merging compact objects (COs) has revolutionized our understanding of both fundamental physics and astronomy~\citep[e.g.][]{LIGOScientific:2021djp,LIGOScientific:2021sio,LIGOScientific:2021psn}.
Although the data volumes from GW detectors are relatively small, analyzing the data is a computationally demanding task.
In addition, this computational cost will substantially increase when next generation detectors come online~\citep{Maggiore:2019uih, Reitze:2019iox, Evans:2021gyd}.

@@

Derivatives are ubiquitously useful throughout data analysis tasks.
For instance, during PE, derivative information can be used to guide the sampling towards regions with higher likelihood values \dfm{This isn't correct. In HMC the gradients don't have the behavior. I'd just say something like ``gradients are used to improve performance''} (e.g.~in Hamiltonian Monte Carlo~\citep{2017arXiv170102434B} or Gradient Descent~\citep{2016arXiv160904747R}).
Gradients are particularly valuable for high dimensional spaces.
Unfortunately, in the field of GW data analysis, analytic derivatives of the necessary quantities (such as the likelihood) have historically been difficult to obtain~\citep{Keppel:2013kia}.
Numerical derivatives also suffer from accuracy issues stemming from rounding or truncation errors.
However, recent progress in automatic differentiation (AD) has shown promise in allowing general, fast derivative calcuations for gravitational waveforms~\citep{Coogan:2022qxs}.

>>>>

The discovery of gravitational waves (GWs)~\citep{LIGOScientific:2016aoc} from inspiraling and merging compact objects (COs) has revolutionized our understanding of both fundamental physics and astronomy~\citep[e.g.][]{LIGOScientific:2021djp,LIGOScientific:2021sio,LIGOScientific:2021psn}.
Although the data volumes from GW detectors are relatively small, analyzing the data is a computationally demanding task.
In addition, this computational cost will substantially increase when next generation detectors come online~\citep{Maggiore:2019uih, Reitze:2019iox, Evans:2021gyd}.

@@

Derivatives are ubiquitously useful throughout data analysis tasks.
For instance, during PE, derivative information can be used to guide the sampling towards regions with higher likelihood values \dfm{This isn't correct. In HMC the gradients don't have the behavior. I'd just say something like ``gradients are used to improve performance''} (e.g.~in Hamiltonian Monte Carlo~\citep{2017arXiv170102434B} or Gradient Descent~\citep{2016arXiv160904747R}).
Gradients are particularly valuable for high dimensional spaces.
Unfortunately, in the field of GW data analysis, analytic derivatives of the necessary quantities (such as the likelihood) have historically been difficult to obtain~\citep{Keppel:2013kia}.
Numerical derivatives also suffer from accuracy issues stemming from rounding or truncation errors.
However, recent progress in automatic differentiation (AD) has shown promise in allowing general, fast derivative calculations for gravitational waveforms\AZ{, with applications to constructing template banks~\citep{Coogan:2022qxs} and computing the Fisher information for forecasting~\citep{Iacovelli:2022bbs, Iacovelli:2022mbg}}.
